Immutable Source-Mapped Abstract Syntax Tree:\ \, \\ A Design Pattern for Refactoring Engine APIs

### 2017-10-04 paper.tex
Almost final touches

There are two main variants of abstract syntax tree:
mutable and immutable. A mutable tree has nodes that can be easily added, 
changed, or deleted from the tree. 
These trees are often used internally by

\begin{wrapfigure}{r}{3in} % r is right, R is float right
% Original image is 4.624 x 10.651 in
\includegraphics[scale=0.62]{jdt-ast.eps}
%\caption{Visualization of the Eclipse Java Development Tools' abstract syntax
%tree for the following program: \\
%\ \\
%\centerline{\parbox{2in}{\ttt{\noindent public class Example \{ \\
%\hspace*{.2em} public int FIVE = 5; \\
%\}
%}}} \\
%\ \\
%The highlighted node---a \textit{VariableDeclarationFragment}---corresponds to
%the eight-character substring \ttt{FIVE = 5}, which begins with the 36th
%character in the source code.
%}
\caption{Visualization of an AST produced by the Eclipse Java Development Tools
(JDT)\vspace*{-4em}}
\label{fig:jdt-ast}
\end{wrapfigure}

compilers. For example, a compiler might determine that
\ttt{a~=~b*1} can be simplified to \ttt{a~=~b}, and then change

the AST accordingly. Immutable trees, on the other hand, cannot be changed via
function calls. These can be used for read-only purposes or operated on in a
functional manner.

@@

\subsection{Basic Forces}

There are several basic requirements that a solution API must meet.

@@

\subsection{AST Forces}
\label{ss:manip}

Any candidate solution would have to satisfy the API requirements while avoiding
the problems associated with mutable ASTs.

@@

\clearpage
\section{Solution}

@@

\subsection{Fulfilling the Requirements of Refactoring}
\label{ss:reqs}

>>>>

There are two main variants of abstract syntax tree:
mutable and immutable. A mutable tree has nodes that can be easily added, 
changed, or deleted from the tree. 
These trees are often used internally by compilers.
\begin{wrapfigure}{r}{3in} % r is right, R is float right
% Original image is 4.624 x 10.651 in
\includegraphics[scale=0.62]{jdt-ast.eps}
%\caption{Visualization of the Eclipse Java Development Tools' abstract syntax
%tree for the following program: \\
%\ \\
%\centerline{\parbox{2in}{\ttt{\noindent public class Example \{ \\
%\hspace*{.2em} public int FIVE = 5; \\
%\}
%}}} \\
%\ \\
%The highlighted node---a \textit{VariableDeclarationFragment}---corresponds to
%the eight-character substring \ttt{FIVE = 5}, which begins with the 36th
%character in the source code.
%}
\caption{Visualization of an AST produced by the Eclipse Java Development Tools
(JDT)\vspace*{-4em}}
\label{fig:jdt-ast}
\end{wrapfigure}
For example, a compiler might determine that
\ttt{a~=~b*1} can be simplified to \ttt{a~=~b}, and then change 
the AST accordingly. Immutable trees, on the other hand, cannot be changed via
function calls. These can be used for read-only purposes or operated on in a
functional manner.

@@

\label{ss:reqs}
\subsection{Basic Forces}

To be useful for implementing refactorings, and API exposed by a refactoring 
tool must meet several basic requirements:

@@

\subsection{Scalability Forces}

As a code base increases in size, so does its complexity. A refactoring API 
must be able to handle the constraints that a more complicated program 
introduces:

\begin{itemize}
    \item \textit{The API should describe the program's structure at a specific 
        point in time, either before or after the a refactoring transformation.} 
        It should never be inconsistent and provide some information that is
        up-to-date with the latest transformation while providing some
        informations that isn't.
\end{itemize}
\clearpage
\begin{itemize}
    \item \textit{The API should be robust in the presence of attempted refactorings
        that produce invalid code.} The refactoring should never leave a 
        client with code that doesn't compile. 
    \item \textit{The API should be implemented in a way that allows easy analysis of
        the syntactic effects of any given change.} This implementation should
        be memory and time efficient, allowing clients to write refactorings
        that meet the performance requirements of an interactive tool.
    \item \textit{The API should be robust in the presence of "malicious" clients.}
        its implementation should not allow clients to manipulate or 
        invalidate internal data structures. This requires an understanding of
        possible states of those data structures. In particular, an AST used
        as an internal data structure might be exposed differently to clients
        via the API. 
\end{itemize}

@@

\section{Solution}

@@

\subsection{Fulfilling the Requirements of Refactoring}
